---
title: "Peer Graded"
author: "Redson"
date: "2023-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Loading the packages required**
```{r}
library(data.table)
library(timeDate)
library(tidyverse)
library(lattice)
```

**Loading the data**
```{r}
data<-read.csv("./activity.csv")
```

**Exploring the data**
```{r}
summary(data)
head(data)
```

# Step 1: Histogram of the total number of steps taken each day
```{r}
total_steps_per_day <- data %>%
  group_by(date) %>%
  summarise(total_steps = sum(steps, na.rm = TRUE))
```

```{r}
# Plot histogram
hist(total_steps_per_day$total_steps, main = "Histogram of Total Steps per Day", xlab = "Total Steps")
```

```{r}
# Step 2: Mean and median number of steps taken each day
mean_steps_per_day <- mean(total_steps_per_day$total_steps, na.rm = TRUE)
median_steps_per_day <- median(total_steps_per_day$total_steps, na.rm = TRUE)
```

```{r}
# Print mean and median
cat("Mean steps per day:", mean_steps_per_day, "\n")
cat("Median steps per day:", median_steps_per_day, "\n")
```
# Step 3: Time series plot of the average number of steps taken
```{r}
average_steps_per_interval <- data %>%
  group_by(interval) %>%
  summarise(average_steps = mean(steps, na.rm = TRUE))
```

# Plot time series
```{r}
plot(average_steps_per_interval$interval, average_steps_per_interval$average_steps, type = "l", 
     main = "Time Series Plot of Average Steps per Interval", xlab = "Interval", ylab = "Average Steps")
```

# Step 4: The 5-minute interval that, on average, contains the maximum number of steps
```{r}
max_interval <- average_steps_per_interval$interval[which.max(average_steps_per_interval$average_steps)]
cat("Interval with the maximum average steps:", max_interval, "\n")
```
# Step 5: Code to describe and show a strategy for imputing missing data
```{r}
mean_steps <- mean(data$steps, na.rm = TRUE)
data$steps[is.na(data$steps)] <- mean_steps
```

# Step 6: Histogram of the total number of steps taken each day after missing values are imputed
```{r}
total_steps_per_day_imputed <- data %>%
  group_by(date) %>%
  summarise(total_steps = sum(steps))
```

# Plot histogram
```{r}
hist(total_steps_per_day_imputed$total_steps, main = "Histogram of Total Steps per Day (Imputed)", xlab = "Total Steps")
```

# Step 7: Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
data$day <- weekdays(as.Date(data$date))
data$day_type <- ifelse(data$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")

average_steps_by_day_type <- data %>%
  group_by(interval, day_type) %>%
  summarise(average_steps = mean(steps))
```
# Plot panel plot
```{r}
ggplot(average_steps_by_day_type, aes(x = interval, y = average_steps, color = day_type)) +
  geom_line() +
  labs(title = "Average Steps per 5-Minute Interval Across Weekdays and Weekends", x = "Interval", y = "Average Steps") +
  theme_minimal()
```

